`flippant` -- An `R` package for the automated  analysis of dithionite scramblase

### 2015-09-19 inst/manuscript/manuscript.rmd
Various typo fixes;

Based on @menon_opsin_2011 and @goren_constitutive_2014, data is processed as 
follows. Input is format checked and defaults are injected for facultative 
parameters/columns as appropriate. Fluorescence spectra are parsed. This includes 
algorithmic determination of when dithionite was added to the sample using peak
detection by continuous wavelet transformation as implemented in the 
`wmtsa` `R` package [@percival_wavelet_2000;@constantine_wmtsa:_2013]. 
Acquisition time is aligned for all spectra in the data set such that the zero 
time point henceforth corresponds to the time of addition. 
Pre-dithionite-addition baseline fluorescence $F_{\text{baseline}}$ is 
determined for each spectrum by averaging (median) over the ten values preceding
dithionite addition. Post-dithinonite-addition minimum fluorescence 
$F_{\text{min}}$ is analogously calculated from the last ten datapoints after 
a default 400 s or the timepoint of measurement supplied by the user in the 
parameter spreadsheet. $F_{\text{min}}$ is volume-corrected based on supplied 
reaction volumes with and without dithionite and for each spectrum/datapoint a 
fluorescence reduction is calculated as follows:
$$
y_{\text{raw}} = 1 - \frac{F_{\text{min}}}{F_{\text{baseline}}}
$$
A relative fluorescence reduction $y$ is calculated scaled to a (required) liposomes-only/no-protein control and the probability $P_{\geq 1}$ for a 
liposome to contain at least one scramblase molecules is calculated using
$$
P_{\geq 1} = \frac{(y - y_0)}{(y_{\text{max}} - y_0)}
$$
where $y$ is the relative fluorescence reduction for the proteoliposome 
measurement and $y_0$ for a measurement of protein-free liposomes. 
$y_{\text{max}}$ refers to either the maximal $y$ in the experiment or is 
derived from a monoexponential fit to the data (default) as a precaution for the 
case where the protein/phospholipid titration did not reach the plateau of the 
saturation curve. The behavior is adjustable using the `scaleTo` parameter of 
the corresponding functions (see below). Next a monoexponential curve is fitted
to the realtionship of $P_{\geq 1}$ to the protein to phospholipid ratio (PPR; 
(mg/mmol)), providing the scramblase-characterizing fit factor $\alpha$.

#Main Features
##Template Generation for Data Input
Supporting the user in providing the input data needed, the following function 
generates a spreadsheet program-compatible tab-delimited `ASCII` template, 
including column names, commentary, expected data type and the default used when
ommited.
```{r, code=formatR::usage('scramblaseAssayInputTemplate'), eval=FALSE}
```
The application agnostic tab-delimited format must be maintained for the 
completed table, which is subjected to extensive input checks when read by 
`flippant`.

##Graphic Representation of Spectral Traces
In a first, graphical data analytics step of a scramblase assay-derived data 
set, fluorescent spectra or traces may be plotted in publication-appropriate
quality with acquisition time in the x- and fluorescense intensity on the y-axis
using
```{r, code = formatR::usage('scramblaseAssayTraces'), eval = FALSE}
```
and handing the function the path to the user assembled spreadsheet of input 
data (`x`). The `adjust` argument to the function call serves to selectively 
inactivate `wmtsa`-provided (see above) alignment of multiple spectra to the 
time of dithionite addition. Acquisition time coverage of the plot may be tuned 
by user-provided minimal (`timeMin` ) and maximal (`timeMax`) time points to be
included.

##Scramblase Assay Analysis
Plotting protein to phospholipid ratio (PPR) against the likelyhood for a given
proteoliposome to contain one or more scramblases, the PPR plot is at the heart 
of the analysis of dithionite scramblase assays, allowing for the visual 
comparison of specific scramblase activity between interrogated experimental 
series, which may e.g. be representative of protein preparations including 
mutant protein forms or crude extracts with or without depletion of a candidate
scramblase. 
```{r, code = formatR::usage('scramblaseAssayPlot'), eval = FALSE}
```
allows for the complete generation of the PPR plot in publication quality from 
the user assembled spreadsheet of input data (`x`), including automatically 
added graphical representations of $\tau$, the PPR value equal to the fit 
constant $\alpha$, which implies all liposomes on average carrying one 
scramblase and 63% have one or more (i.e. are active). The behavior of the 
function may be adjusted to use scramblase-mediated fluorescense reduction 
normalized to either the maximum measurement in the series (`scaleTo = "data"`)
or the plateau of an exponential fit to it (`scaleTo = "model"`; for reasoning
see above) and whether or not to require the exponential curve fitted to the 
data to go through the origin.

@@

#Funding
J.G., R.J.C. and the Proteomics Core at WCMC-Q were supported by ``Biomedical 
Research Program'' funds at Weill Cornell Medical College in Qatar, a program 
funded by Qatar Foundation. B.P. was supported by a grant from Qatar National 
Research Fund’s National Priority Research Program (5-669-1-112) to J.G. and
A.K.M.

>>>>

Based on @menon_opsin_2011 and @goren_constitutive_2014, data is processed as 
follows. Input is format checked and defaults are injected for facultative 
parameters/columns as appropriate. Fluorescence spectra are parsed. This includes 
algorithmic determination of when dithionite was added to the sample using peak
detection by continuous wavelet transformation as implemented in the 
`wmtsa` `R` package [@percival_wavelet_2000;@constantine_wmtsa:_2013]. 
Acquisition time is aligned for all spectra in the data set such that the zero 
time point henceforth corresponds to the time of addition. 
Pre-dithionite-addition baseline fluorescence $F_{\text{baseline}}$ is 
determined for each spectrum by averaging (median) over the ten values preceding
dithionite addition. Post-dithionite-addition minimum fluorescence 
$F_{\text{min}}$ is analogously calculated from the last ten data points after 
a default 400 s or the time point of measurement supplied by the user in the 
parameter spreadsheet. $F_{\text{min}}$ is volume-corrected based on supplied 
reaction volumes with and without dithionite and for each spectrum/data point a 
fluorescence reduction is calculated as follows:
$$
y_{\text{raw}} = 1 - \frac{F_{\text{min}}}{F_{\text{baseline}}}
$$
A relative fluorescence reduction $y$ is calculated scaled to a (required) liposomes-only/no-protein control and the probability $P_{\geq 1}$ for a 
liposome to contain at least one scramblase molecules is calculated using
$$
P_{\geq 1} = \frac{(y - y_0)}{(y_{\text{max}} - y_0)}
$$
where $y$ is the relative fluorescence reduction for the proteoliposome 
measurement and $y_0$ for a measurement of protein-free liposomes. 
$y_{\text{max}}$ refers to either the maximal $y$ in the experiment or is 
derived from a monoexponential fit to the data (default) as a precaution for the 
case where the protein/phospholipid titration did not reach the plateau of the 
saturation curve. The behavior is adjustable using the `scaleTo` parameter of 
the corresponding functions (see below). Next a monoexponential curve is fitted
to the relationship of $P_{\geq 1}$ to the protein to phospholipid ratio (PPR; 
(mg/mmol)), providing the scramblase-characterizing fit factor $\alpha$.

#Main Features
##Template Generation for Data Input
Supporting the user in providing the input data needed, the following function 
generates a spreadsheet program-compatible tab-delimited `ASCII` template, 
including column names, commentary, expected data type and the default used when
omitted.
```{r, code=formatR::usage('scramblaseAssayInputTemplate'), eval=FALSE}
```
The application agnostic tab-delimited format must be maintained for the 
completed table, which is subjected to extensive input checks when read by 
`flippant`.

##Graphic Representation of Spectral Traces
In a first, graphical data analytic step of a scramblase assay-derived data 
set, fluorescent spectra or traces may be plotted in publication-appropriate
quality with acquisition time in the x- and fluorescence intensity on the y-axis
using
```{r, code = formatR::usage('scramblaseAssayTraces'), eval = FALSE}
```
and handing the function the path to the user assembled spreadsheet of input 
data (`x`). The `adjust` argument to the function call serves to selectively 
inactivate `wmtsa`-provided (see above) alignment of multiple spectra to the 
time of dithionite addition. Acquisition time coverage of the plot may be tuned 
by user-provided minimal (`timeMin` ) and maximal (`timeMax`) time points to be
included.

##Scramblase Assay Analysis
Plotting protein to phospholipid ratio (PPR) against the likelihood for a given
proteoliposome to contain one or more scramblases, the PPR plot is at the heart 
of the analysis of dithionite scramblase assays, allowing for the visual 
comparison of specific scramblase activity between interrogated experimental 
series, which may e.g. be representative of protein preparations including 
mutant protein forms or crude extracts with or without depletion of a candidate
scramblase. 
```{r, code = formatR::usage('scramblaseAssayPlot'), eval = FALSE}
```
allows for the complete generation of the PPR plot in publication quality from 
the user assembled spreadsheet of input data (`x`), including automatically 
added graphical representations of $\tau$, the PPR value equal to the fit 
constant $\alpha$, which implies all liposomes on average carrying one 
scramblase and 63% have one or more (i.e. are active). The behavior of the 
function may be adjusted to use scramblase-mediated fluorescence reduction 
normalized to either the maximum measurement in the series (`scaleTo = "data"`)
or the plateau of an exponential fit to it (`scaleTo = "model"`; for reasoning
see above) and whether or not to require the exponential curve fitted to the 
data to go through the origin.

@@

#Funding
J.G., R.J.C. and the Proteomics Core at WCMC-Q are supported by ``Biomedical 
Research Program'' funds at Weill Cornell Medical College in Qatar, a program 
funded by Qatar Foundation. B.P. was supported by a grant from Qatar National 
Research Fund’s National Priority Research Program (5-669-1-112) to J.G. and
A.K.M.
